Click chemistry reveals an alternate mechanism for HPII catalase

### 2018-02-05 bb494/manuscript.tex
more

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Meta-data block
%% ---------------
%% Each author should be given as a separate \author command.
%%
%% Corresponding authors should have an e-mail given after the author
%% name as an \email command. Phone and fax numbers can be given
%% using \phone and \fax, respectively; this information is optional.
%%
%% The affiliation of authors is given after the authors; each
%% \affiliation command applies to all preceding authors not already
%% assigned an affiliation.
%%
%% The affiliation takes an option argument for the short name.  This
%% will typically be something like "University of Somewhere".
%%
%% The \altaffiliation macro should be used for new address, etc.
%% On the other hand, \alsoaffiliation is used on a per author basis
%% when authors are associated with multiple institutions.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\author{Elliott Capek}
\affiliation{Department of Biochemistry, Oregon State University}
\email{capeke@oregonstate.edu}
\phone{(971) 207 8193}
%% \alsoaffiliation[]{}
\author{Aidan Estelle}
%% \affiliation{Department of Biochemistry, Oregon State University}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Start the main part of the manuscript here.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
Catalase is an enzyme which catalyzes the reaction $2H_2O_2 \rightarrow O_2 + 2H_2O$. This is an important detoxification reaction, since hydrogen peroxide is a reactive toxin generated by several biological processes.

@@

This bimolecular mechanism relies on the efficient transport of hydrogen peroxide into the deeply buried active site. This is believed to be aided by an electrical gradient over 12 \AA between an active site aspartic acid and the positive heme iron \cite{chelikani}. Mutagenisis of this Asp to positive residues decreases activity up to 90\% \cite{chelikani}. However, this study was accomplished using canonical amino acids, and provides relatively weak evidence for the electric field hypothesis.\\

In this study we seek to use the technique of bioorthogonal ligation to study the electric field hypothesis of catalase substrate transport. We will incorporate the noncanonical amino acid 3-nitrotyrosine into the active site, then use click chemistry to insert a gradient of charged particles into the active site. By using a large number of particles with charges varying from positive to neutral to negative, we can average out the effects of steric bulk and isolate the electrical effects of charged amino acids in the catalase active site. The results of such a study would conclusively determine the importance of an electric field in the catalase active site.\\

@@

\section{Results}
DH10B cell lines with appropriate plasmids were grown for 48h and pelletized. Wildtype HPII-containing cells were purified using TALON resin. Purified wildtype protein and ncAA cell lysates are shown in Figure \ref{fig:crude-wt-gel}. The expected band corresponding to an 84-kD protein is seen across all lanes, verifying the presence of HPII wildtype in all cultures.\\

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Meta-data block
%% ---------------
%% Each author should be given as a separate \author command.
%%
%% Corresponding authors should have an e-mail given after the author
%% name as an \email command. Phone and fax numbers can be given
%% using \phone and \fax, respectively; this information is optional.
%%
%% The affiliation of authors is given after the authors; each
%% \affiliation command applies to all preceding authors not already
%% assigned an affiliation.
%%
%% The affiliation takes an option argument for the short name.  This
%% will typically be something like "University of Somewhere".
%%
%% The \altaffiliation macro should be used for new address, etc.
%% On the other hand, \alsoaffiliation is used on a per author basis
%% when authors are associated with multiple institutions.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\author{Elliott Capek}
\affiliation{Department of Biochemistry, Oregon State University}
\email{capeke@oregonstate.edu}
\phone{(971) 207 8193}
\author{Hassan Alnatah}
\author{Laikana Ly}
\author{Joey Orton}
%% \alsoaffiliation[]{}
\author{Aidan Estelle}
%% \affiliation{Department of Biochemistry, Oregon State University}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Start the main part of the manuscript here.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
Catalase is an enzyme which catalyzes the reaction $2H_2O_2 \rightarrow O_2 + 2H_2O$. This is an important detoxification reaction, since hydrogen peroxide is a reactive toxin generated by several biological processes. Catalase is a tetrameric protein which uses a heme group in its active site to facilitate the breakdown of hydrogen peroxide. It is a unique protein in that its reaction evolves gas, and thus its activity can be seen with the bare eye. The protein has been studied as a target for tumor suppression drugs \cite{bauer}. The enzyme is also useful industrially, in particular in the bleaching of textiles \cite{amorim}.

@@

This bimolecular substrate mechanism relies on the efficient transport of hydrogen peroxide into the deeply buried active site. This is believed to be aided by an electrical gradient over 12 \AA between an active site aspartic acid and the positive heme iron \cite{chelikani}. Mutagenisis of this Asp to positive residues decreases activity up to 90\% \cite{chelikani}. However, this study was accomplished using only a few canonical amino acid substitutions, and provides relatively weak evidence for the electric field hypothesis.

In this study we seek to use the technique of bioorthogonal ligation to study the electric field hypothesis of catalase substrate transport. We will incorporate the noncanonical amino acid 3-nitrotyrosine into the active site, then use click chemistry to insert a gradient of charged particles into the active site. By using a large number of particles with charges varying from positive to negative, we can average out the effects of steric bulk and isolate the effect of field strength. The results of such a study would conclusively determine the importance of an electric field in the catalase active site.\\

@@

\section{Results}
DH10B cell lines with appropriate plasmids were grown for 48h and pelletized. Wildtype HPII-containing cells were purified using TALON resin. Purified wildtype protein and ncAA cell lysates are shown in Figure \ref{fig:crude-wt-gel}. The expected band corresponding to an 84-kD protein is seen across all lanes, verifying the presence of HPII wildtype in all cultures. Contaminant bands in the gel are very faint, indicating pure protein.\\
